---
title: "Maps"
author: "Agasiya"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "NA", warning = FALSE, message = FALSE)
```

##Introduction

#Installation

```{r}
#install.packages("leaflet")
# to install the development version from Github, run
# devtools::install_github("rstudio/leaflet")
```

```{r}
library(leaflet)
library(maps)
library(sp)
library(htmltools)
```

#Basic Usage
You create a Leaflet map with these basic steps:

Create a map widget by calling leaflet().
Add layers (i.e., features) to the map by using layer functions (e.g. addTiles, addMarkers, addPolygons) to modify the map widget.
Repeat step 2 as desired.
Print the map widget to display it.

```{r}
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng = -1.576121, lat = 6.6865339, popup="Kotei")
m  # Print the map
```



In case you’re not familiar with the magrittr pipe operator (%>%), here is the equivalent without using pipes

```{r}
m <- leaflet()
m <- addTiles(m)
m <- addMarkers(m, lng=-1.576121, lat=6.6865339, popup="Kotei")
m
```

##The Map Widget
The function leaflet() returns a Leaflet map widget, which stores a list of objects that can be modified or updated later. Most functions in this package have an argument map as their first argument, which makes it easy to use the pipe operator %>% in the magrittr package, as you have seen from the example in the Introduction.

#Initializing Options
The map widget can be initialized with certain parameters. This is achieved by populating the options argument as shown below.

```{r}
# Set value for the minZoom and maxZoom settings.
leaflet(options = leafletOptions(minZoom = 0, maxZoom = 18))
```

The leafletOptions() can be passed any option described in the leaflet reference document. Using the leafletOptions(), you can set a custom CRS and have your map displayed in a non spherical mercator projection as described in projections.

##Map Methods
You can manipulate the attributes of the map widget using a series of methods. Please see the help page ?setView for details.

setView() sets the center of the map view and the zoom level;
fitBounds() fits the view into the rectangle [lng1, lat1] – [lng2, lat2];
clearBounds() clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided;

##The Data Object
Both leaflet() and the map layer functions have an optional data parameter that is designed to receive spatial data in one of several forms:

#From base R:
lng/lat matrix
data frame with lng/lat columns
#From the sp package:
SpatialPoints[DataFrame]
Line/Lines
SpatialLines[DataFrame]
Polygon/Polygons
SpatialPolygons[DataFrame]
#From the maps package:
the data frame from returned from map()

The data argument is used to derive spatial data for functions that need it; for example, if data is a SpatialPolygonsDataFrame object, then calling addPolygons on that map widget will know to add the polygons from that SpatialPolygonsDataFrame.

It is straightforward to derive these variables from sp objects since they always represent spatial data in the same way. On the other hand, for a normal matrix or data frame, any numeric column could potentially contain spatial data. So we resort to guessing based on column names:

the latitude variable is guessed by looking for columns named lat or latitude (case-insensitive)
the longitude variable is guessed by looking for lng, long, or longitude
You can always explicitly identify latitude/longitude columns by providing lng and lat arguments to the layer function.

For example, we do not specify the values for the arguments lat and lng in addCircles() below, but the columns Lat and Long in the data frame df will be automatically used:

```{r}
# add some circles to a map
df = data.frame(Lat = 1:10, Long = rnorm(10))
leaflet(df) %>% addCircles()
```

You can also explicitly specify the Lat and Long columns (see below for more info on the ~ syntax):
```{r}
leaflet(df) %>% addCircles(lng = ~Long, lat = ~Lat)
```

```{r}
mapStates = map("Ghana", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```

##The Formula Interface
The arguments of all layer functions can take normal R objects, such as a numeric vector for the lat argument, or a character vector of colors for the color argument. They can also take a one-sided formula, in which case the formula will be evaluated using the data argument as the environment. For example, ~ x means the variable x in the data object, and you can write arbitrary expressions on the right-hand side, e.g., ~ sqrt(x + 1).
```{r}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```


##Using Basemaps
Leaflet supports basemaps using map tiles, popularized by Google Maps and now used by nearly all interactive web maps.

#Default (OpenStreetMap) Tiles
The easiest way to add tiles is by calling addTiles() with no arguments; by default, OpenStreetMap tiles are used.
```{r}
m <- leaflet() %>% setView(lng = -1.5764576, lat = 6.6698666, zoom = 12)
m %>% addTiles()
```

##Third-Party Tiles
Alternatively, many popular free third-party basemaps can be added using the addProviderTiles() function, which is implemented using the leaflet-providers plugin. See here for the complete set.

As a convenience, leaflet also provides a named list of all the third-party tile providers that are supported by the plugin. This enables you to use auto-completion feature of your favorite R IDE (like RStudio) and not have to remember or look up supported tile providers; just type providers$ and choose from one of the options. You can also use names(providers) to view all of the options.

```{r}
m %>% addProviderTiles(providers$Stamen.Toner)
```

```{r}
m %>% addProviderTiles(providers$CartoDB.Positron)
```

```{r}
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```

##Custom Tile URL Template
If you happen to have a custom map tile URL template to use, you can provide it as an argument to addTiles().

#WMS Tiles
You can use addWMSTiles() to add WMS (Web Map Service) tiles. The map below shows the Base Reflectivity (a measure of the intensity of precipitation occurring) using the WMS from the Iowa Environmental Mesonet:

```{r}
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 4) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```

##Combining Tile Layers
You aren’t restricted to using a single basemap on a map; you can stack them by adding multiple tile layers. This generally only makes sense if the front tiles consist of semi transparent tiles, or have an adjusted opacity via the options argument.
```{r}
m %>% addProviderTiles(providers$MtbMap) %>%
  addProviderTiles(providers$Stamen.TonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stamen.TonerLabels)
```

##Markers
Use markers to call out points on the map. Marker locations are expressed in latitude/longitude coordinates, and can either appear as icons or as circles.

#Data sources
Point data for markers can come from a variety of sources:

SpatialPoints or SpatialPointsDataFrame objects (from the sp package)
POINT, sfc_POINT, and sf objects (from the sf package); only X and Y dimensions will be considered
Two-column numeric matrices (first column is longitude, second is latitude)
Data frame with latitude and logitude columns. You can explicitly tell the marker function which columns contain the coordinate data (e.g. addMarkers(lng = ~Longitude, lat = ~Latitude)), or let the function look for columns named lat/latitude and lon/lng/long/longitude (case insensitive).
Simply provide numeric vectors as lng and lat arguments
Note that MULTIPOINT objects from sf are not supported at this time.

#Icon Markers
Icon markers are added using the addMarkers or the addAwesomeMarkers functions. Their default appearance is a dropped pin. As with most layer functions, the popup argument can be used to add a message to be displayed on click, and the label option can be used to display a text label either on hover or statically.
```{r}
data(quakes)

# Show first 20 rows from the `quakes` dataset
leaflet(data = quakes[1:20,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

##Customizing Marker Icons
You can provide custom markers in one of several ways, depending on the scenario. For each of these ways, the icon can be provided as either a URL or as a file path.

For the simple case of applying a single icon to a set of markers, use makeIcon().
```{r}
greenLeafIcon <- makeIcon(
  iconUrl = "https://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

If you have several icons to apply that vary only by a couple of parameters (i.e. they share the same size and anchor points but have different URLs), use the icons() function. icons() performs similarly to data.frame(), in that any arguments that are shorter than the number of markers will be recycled to fit.
```{r}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "https://leafletjs.com/examples/custom-icons/leaf-green.png",
    "https://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "https://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = leafIcons)
```

Finally, if you have a set of icons that vary in multiple parameters, it may be more convenient to use the iconList() function. It lets you create a list of (named or unnamed) makeIcon() icons, and select from that list by position or name.
```{r}
# Make a list of icons. We'll index into it based on name.
oceanIcons <- iconList(
  ship = makeIcon("ferry-18.png", "ferry-18@2x.png", 18, 18),
  pirate = makeIcon("danger-24.png", "danger-24@2x.png", 24, 24)
)

# Some fake data
df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 1.5491641,  # lng
    (runif(20) - .5) * 3.8 + 6.6612312  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "pirate", "ship"),
    c("ship", "pirate")
  ))
)

leaflet(df) %>% addTiles() %>%
  # Select from oceanIcons based on df$type
  addMarkers(icon = ~oceanIcons[type])
```

Awesome Icons
Leaflet supports even more customizable markers using the awesome markers leaflet plugin.

The addAwesomeMarkers() function is similar to addMarkers() function but additionally allows you to specify custom colors for the markers as well as icons from the Font Awesome, Bootstrap Glyphicons, and Ion icons icon libraries.

Similar to the makeIcon, icons, and iconList functions described above, you have makeAwesomeIcon, awesomeIcons and awesomeIconList functions, which enable you to add awesome icons.
```{r}
# first 20 quakes
df.20 <- quakes[1:20,]

getColor <- function(quakes) {
  sapply(quakes$mag, function(mag) {
  if(mag <= 4) {
    "green"
  } else if(mag <= 5) {
    "orange"
  } else {
    "red"
  } })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(df.20)
)

leaflet(df.20) %>% addTiles() %>%
  addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(mag))
```

The library argument has to be one of ‘ion’, ‘fa’, or ‘glyphicon’. The icon argument needs to be the name of any valid icon supported by the the respective library (w/o the prefix of the library name).

#Marker Clusters
When there are a large number of markers on a map, you can cluster them using the Leaflet.markercluster plug-in. To enable this plug-in, you can provide a list of options to the argument clusterOptions, e.g.
```{r}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions()
)
```

Using the freezeAtZoom argument of the markerClusterOptions() function you can set the clustering to freeze as a specific zoom level. For example markerClusterOptions(freezeAtZoom = 5) will freeze the cluster at zoom level 5 regardless of the user’s actual zoom level.

#Circle Markers
Circle markers are much like regular circles (see Lines and Shapes), except that their radius in onscreen pixels stays constant regardless of zoom level.

You can use their default appearance:
```{r}
leaflet(df) %>% addTiles() %>% addCircleMarkers()
```

Or customize their color, radius, stroke, opacity, etc.
```{r}
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )
```

##Popups
Popups are small boxes containing arbitrary HTML, that point to a specific point on the map.

Use the addPopups() function to add standalone popup to the map.
```{r}
content <- paste(sep = "<br/>",
  "<b><a href='http://www.samurainoodle.com'>Samurai Noodle</a></b>",
  "606 5th Ave. S",
  "Seattle, WA 98138"
)

leaflet() %>% addTiles() %>%
  addPopups(-122.327298, 47.597131, content,
    options = popupOptions(closeButton = FALSE)
  )
```

A common use for popups is to have them appear when markers or shapes are clicked. Marker and shape functions in the Leaflet package take a popup argument, where you can pass in HTML to easily attach a simple popup.

```{r}
df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"
))

leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, popup = ~htmlEscape(Name))
```

In the preceding example, htmltools::htmlEscape was used to santize any characters in the name that might be interpreted as HTML. While it wasn’t necessary for this example (as the restaurant names contained no HTML markup), doing so is important in any situation where the data may come from a file or database, or from the user.

In addition to markers you can also add popups on shapes like lines, circles and other polygons.

##Labels
A label is a textual or HTML content that can attached to markers and shapes to be always displayed or displayed on mouse over. Unlike popups you don’t need to click a marker/polygon for the label to be shown.
```{r}
df <- read.csv(textConnection(
"Name,Lat,Long
Samurai Noodle,47.597131,-122.327298
Kukai Ramen,47.6154,-122.327157
Tsukushinbo,47.59987,-122.326726"))

leaflet(df) %>% addTiles() %>%
  addMarkers(~Long, ~Lat, label = ~htmlEscape(Name))
```

##Customizing Marker Labels
You can customize marker labels using the labelOptions argument of the addMarkers function. The labelOptions argument can be populated using the labelOptions() function. If noHide is false (the default) then the label is displayed only when you hover the mouse over the marker; if noHide is set to true then the label is always displayed.
```{r}
# Change Text Size and text Only and also a custom CSS
leaflet() %>% addTiles() %>% setView(-118.456554, 34.09, 13) %>%
  addMarkers(
    lng = -118.456554, lat = 34.105,
    label = "Default Label",
    labelOptions = labelOptions(noHide = T)) %>%
  addMarkers(
    lng = -118.456554, lat = 34.095,
    label = "Label w/o surrounding box",
    labelOptions = labelOptions(noHide = T, textOnly = TRUE)) %>%
  addMarkers(
    lng = -118.456554, lat = 34.085,
    label = "label w/ textsize 15px",
    labelOptions = labelOptions(noHide = T, textsize = "15px")) %>%
  addMarkers(
    lng = -118.456554, lat = 34.075,
    label = "Label w/ custom CSS style",
    labelOptions = labelOptions(noHide = T, direction = "bottom",
      style = list(
        "color" = "red",
        "font-family" = "serif",
        "font-style" = "italic",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"
      )))
```


##Circles
Circles are added using addCircles(). Circles are similar to circle markers; the only difference is that circles have their radii specified in meters, while circle markers are specified in pixels. As a result, circles are scaled with the map as the user zooms in and out, while circle markers remain a constant size on the screen regardless of zoom level.

When plotting circles, only the circle centers (and radii) are required, so the set of valid data sources is different than for polygons and the same as for markers. See the introduction to Markers for specifics.
```{r}
cities <- read.csv(textConnection("
City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994
"))

leaflet(cities) %>% addTiles() %>%
  addCircles(lng = ~Long, lat = ~Lat, weight = 1,
    radius = ~sqrt(Pop) * 30, popup = ~City
  )
```

##Rectangles
Rectangles are added using the addRectangles() function. It takes lng1, lng2, lat1, and lat2 vector arguments that define the corners of the rectangles. These arguments are always required; the rectangle geometry cannot be inferred from the data object.
```{r}
leaflet() %>% addTiles() %>%
  addRectangles(
    lng1=-118.456554, lat1=34.078039,
    lng2=-118.436383, lat2=34.062717,
    fillColor = "transparent"
  )
```

